Capon Beamforming and Moving Objects - An Analysis of Lateral Shift-Invariance

### 2014-04-24 HRUIofMovingObjects/response_to_reviewers.tex

\\\\
Equation (8) the two factors p and q are confusing. Because q is always one, I would suggest omitting it.  Let us say that the oversampling is used both for Capon and for reducing folding of frequencies over the bandwidth.
\\\\
\textit{We understand that this may be confusing and we have amended the text to make it more clear ($q$ is e.g. 4 and 16 when oversampling is applied for Capon beamforming). The factor $p$ has been removed, and $q$ is now the oversampling factor. An additional 2-factor has been added to Eq. (8) with a new reference to sampling of two-way imaging systems ([21]). This factor of two replaces the previous use of $p$.}

@@

\\\\
At what depth are the LSV plots taken with respect to the focal depth? Does it matter?
\\\\
\textit{Yes. Shift invariance caused by under sampling will be at its maximum at the transmit focus (See [22]) (the two-way beam is at its narrowest at this location). Simulated LSV-plots are generated from this point, as specified on page 5, first paragraph. In the in-vitro data, the point scatterer closest to the transmit focus is selected. We have amended the text to clearly state why the transmit focus has been selected as the point of observation.}

### 2014-04-24 HRUIofMovingObjects/jrnl_article.tex
Background. GPU programming.

\section{In-vitro phantom data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in-vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S-D phased transducer operating in harmonic mode at 3.34 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency, and 111 tx-beams where therefore distributed on a 65\degree sector. Transmit focus is located at 8 cm range. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:phantom_das_image} shows the first frame out of 14 in the acquired loop processed with DAS beamforming. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

@@

%\subsection{In-vitro Images}
%Acquire channel data from a phantom containing point wire targets and cysts while moving the probe laterally. Confirm the improved shift-invariance when using phase rotation.

\begin{figure*}[!t]
\centerline{
\begin{tabular}{cc}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6a.png}\label{fig:phantom_das_image}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6b.eps}\label{fig:phantom_das_LSV}}\\
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6c.eps}\label{fig:phantom_capon_LSV}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6d.eps}\label{fig:phantom_capon_LSV_16PR}}
\end{tabular}
}
\caption{LSV-plots generated from \textit{in-vitro} phantom data where the probe has been manually dragged across the phantom. See the supplementary material for videos of the whole loop processed with the different beamformers. Capon beamforming has been performed with the same settings as in Fig.\,\ref{fig:capon1}. P-MAE is not shown due to a lack of probe movement control. a) First frame in the recorded loop processed with delay-and-sum beamforming  \multimedia{Media Movie 1}. A white box indicates the point scatterer used to construct the LSV-plots.  b) LSV-plot for delay-and-sum beamforming. c) LSV-plot for Capon beamforming without oversampling. d) LSV-plot for Capon beamforming with sixteen times oversampling using phase rotation.}
\label{fig:LSV_invitro}
\end{figure*}

@@

Faced with an imaging scene containing bright point scatterers or sharp edges, one could wonder how large the oversampling factor needs to be in order to avoid visible scalloping losses. In his seminal paper, Cox \cite{Cox1973} derives the required beam density for a given output SNR. For instance 47 dB requires a beam density for the Capon beamformer ten times the one given by (\ref{eq:resolution}) [with $q=1/2$]. Since resolution for the Capon beamformer depends on SNR we would recommend to first use the dynamic range in the image as guidance. The amount of diagonal loading will also dictate the maximal achievable resolution since $10\log{d}$ can be interpreted as the maximal input SNR seen by the Capon beamformer. Fig.\,\ref{fig:das_capon_beams} shows that 25 times oversampling is required for a $d$-value of $1/100$. When \textit{in-vitro} data contains additional noise this oversampling requirement is reduced.%Cox measures resolution as the 3 dB saddle point between two impinging monochromatic plane waves. For ultrasound data only a 50 \% increase in resolution was found in \cite[Fig.~7]{Asen} for a 30 dB output SNR when measured in a similar way as Cox [6 dB saddle point]. However, sixteen times oversampling was required to make the imaging system shift invariant in the same paper. In this paper more than ten times oversampling was found to be required for a 40 dB output SNR (after array gain) in simulations, however for the \textit{in-vitro} data ten times oversampling was enough for the system to be locally laterally shift-invariant. This might be a result of improved initial sampling (as discussed in Section \ref{sec:res}) and a higher initial noise level before diagonal loading was applied.

@@

\section{Conclusion}\label{sec:con}
%In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with only a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in-vitro} data when the ultrasound probe is subjected to motion.

In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. It was shown that Capon beamforming with the same beam density as required for delay-and-sum beforming results in blinking point scatterers when the system is subjected to motion.  We then discussed and investigated different methods based on oversampling on transmit and receive in order to reduced this effect. Finally we showed that the local lateral shift-invariance could be improved by oversampling based on phase rotation on receive. This without affecting the acquisition frame rate and with a minor change in processing complexity. In simulations, this method reduced the maximum scalloping loss by 16 dB. On \textit{in-vitro} data the reduction was around 7 dB and the point scatterer blinking was highly reduced.

@@

The authors would like to thank Anders S\o{}rnes at GE Vingmed Ultrasound, Horten, Norway for the help with acquiring \textit{in-vitro} data.

>>>>

\section{In vitro phantom data}\label{sec:res}
In Fig.\,\ref{fig:LSV_invitro} we present the result of applying the Capon beamformer on \textit{in vitro} data with and without oversampling by phase rotation. The data was acquired using a modified Vivid E9 high-end cardiac ultrasound scanner by GE Vingmed Ultrasound, Horten, Norway, equipped with a M5S-D phased transducer operating in harmonic mode at 3.34 MHz. The transmit beam density was set to match (\ref{eq:resolution}) based on this frequency, and 111 tx-beams where therefore distributed on a 65\degree sector. Transmit focus is located at 8 cm range. Note that in harmonic imaging the generated transmit beam at the harmonic frequency will be wider and have lower sidelobes than the beam resulting from transmitting at the harmonic frequency \cite{Fedewa2004}. The lateral sampling is therefore increased compared with our simulations. A tissue mimicking phantom containing wire targets and anechoic cysts was imaged, and the data was acquired while the probe was moved across the phantom manually. Fig.\,\ref{fig:phantom_das_image} shows the first frame out of 14 in the acquired loop processed with DAS beamforming. %Movies of the whole loop processed with DAS and Capon beamforming with and without ten times oversampling by phase rotation can be found in the supplementary material \multimedia{Videos}.

@@

%\subsection{In vitro Images}
%Acquire channel data from a phantom containing point wire targets and cysts while moving the probe laterally. Confirm the improved shift-invariance when using phase rotation.

\begin{figure*}[!t]
\centerline{
\begin{tabular}{cc}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6a.png}\label{fig:phantom_das_image}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6b.eps}\label{fig:phantom_das_LSV}}\\
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6c.eps}\label{fig:phantom_capon_LSV}}
\subfloat[]{\includegraphics[width=0.45\linewidth]{\img fig6d.eps}\label{fig:phantom_capon_LSV_16PR}}
\end{tabular}
}
\caption{LSV-plots generated from \textit{in vitro} phantom data where the probe has been manually dragged across the phantom. See the supplementary material for videos of the whole loop processed with the different beamformers. Capon beamforming has been performed with the same settings as in Fig.\,\ref{fig:capon1}. P-MAE is not shown due to a lack of probe movement control. a) First frame in the recorded loop processed with delay-and-sum beamforming  \multimedia{Media Movie 1}. A white box indicates the point scatterer used to construct the LSV-plots.  b) LSV-plot for delay-and-sum beamforming. c) LSV-plot for Capon beamforming without oversampling. d) LSV-plot for Capon beamforming with sixteen times oversampling using phase rotation.}
\label{fig:LSV_invitro}
\end{figure*}

@@

Faced with an imaging scene containing bright point scatterers or sharp edges, one could wonder how large the oversampling factor needs to be in order to avoid visible scalloping losses. In his seminal paper, Cox \cite{Cox1973} derives the required beam density for a given output SNR. For instance 47 dB requires a beam density for the Capon beamformer ten times the one given by (\ref{eq:resolution}) [with $q=1/2$]. Since resolution for the Capon beamformer depends on SNR we would recommend to first use the dynamic range in the image as guidance. The amount of diagonal loading will also dictate the maximal achievable resolution since $10\log{d}$ can be interpreted as the maximal input SNR seen by the Capon beamformer. Fig.\,\ref{fig:das_capon_beams} shows that 25 times oversampling is required for a $d$-value of $1/100$. When \textit{in vitro} data contains additional noise this oversampling requirement is reduced.%Cox measures resolution as the 3 dB saddle point between two impinging monochromatic plane waves. For ultrasound data only a 50 \% increase in resolution was found in \cite[Fig.~7]{Asen} for a 30 dB output SNR when measured in a similar way as Cox [6 dB saddle point]. However, sixteen times oversampling was required to make the imaging system shift invariant in the same paper. In this paper more than ten times oversampling was found to be required for a 40 dB output SNR (after array gain) in simulations, however for the \textit{in vitro} data ten times oversampling was enough for the system to be locally laterally shift-invariant. This might be a result of improved initial sampling (as discussed in Section \ref{sec:res}) and a higher initial noise level before diagonal loading was applied.

@@

\section{Conclusion}\label{sec:con}
%In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. We have shown that the Capon beamformer is less sensitive to small lateral shifts if oversampling based on phase rotation is applied. This without reducing the acquisition frame rate and with only a minor increase in processing complexity. %In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be too coarse lateral sampling. 
%In Section \ref{sec:methods} we presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity combined, was based on phase rotation. 
%In Section \ref{sec:res} we presented results showing that this method also works on \textit{in vitro} data when the ultrasound probe is subjected to motion.

In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. It was shown that Capon beamforming with the same beam density as required for delay-and-sum beforming results in blinking point scatterers when the system is subjected to motion.  We then discussed and investigated different methods based on oversampling on transmit and receive in order to reduced this effect. Finally we showed that the local lateral shift-invariance could be improved by oversampling based on phase rotation on receive. This without affecting the acquisition frame rate and with a minor change in processing complexity. In simulations, this method reduced the maximum scalloping loss by 16 dB. On \textit{in vitro} data the reduction was around 7 dB and the point scatterer blinking was highly reduced.

@@

The authors would like to thank Anders S\o{}rnes at GE Vingmed Ultrasound, Horten, Norway for the help with acquiring \textit{in vitro} data.
